---
title: "Course Project 2"
author: "Me"
date: "12/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries
```{r}
library(tidyverse)
library(janitor)
library(data.table)
library(lubridate)
library(stringr)
```

Load datasets
```{r}
NEI <- as_tibble(readRDS("summarySCC_PM25.rds")) 
SCC <- as_tibble(readRDS("Source_Classification_Code.rds"))
```

Check data

NEI
```{r}
head(NEI)
dim(NEI)
#no missing value
map_dbl(NEI, function(x) sum(is.na(x)))
```

```{r}
head(SCC)
dim(SCC)
```

SCC is the connection for the NEI. Try to merge

```{r}
all_data = NEI %>%
      left_join(SCC, by = 'SCC')

head(all_data)
dim(all_data)
```

Clean names to make them same
```{r}
all_data = clean_names(all_data)
SCC = clean_names(SCC)
NEI = clean_names(NEI)
```


#1
Have total emissions decreased from 1999 to 2008? Use base plotting system. Make a plot showing total PM2.5 emissions for each years 1999, 2002, 2005, 2008

```{r}
year_total = NEI %>%
      mutate(year = factor(year)) %>%
      group_by(year) %>%
      summarise(total_emission = sum(emissions, na.rm = T))

with(year_total, barplot(total_emission, names = year, ylab = 'total PM2.5 emissions', main = 'Total Annual Emission from 1999 to 2008'))
```


#2. 
Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (fips = '24510') from 1999 to 2008? Use base plotting system

```{r}
baltimore_total_yearly = NEI %>%
      filter(fips == '24510') %>%
      group_by(year) %>%
      summarise(total_emission = sum(emissions, na.rm = TRUE))

with(baltimore_total_yearly, barplot(total_emission, names = year, ylab = 'total PM2.5 emissions', main = 'Total Annual Emission from 1999 to 2008 in Baltimore City'))
```
#3
Of the four types of sources indicated by type variable, which of the four decreases emission from 1999-2008 for baltimore city? Which increased? USe ggplot2

```{r}
graph3 = NEI %>%
      filter(fips == '24510') %>%
      group_by(year, type) %>%
      summarise(total_emission = sum(emissions, na.rm = TRUE)) %>%
      mutate(type = factor(type),
             year = factor(year)) %>%
      ggplot(aes(year, total_emission)) +
      geom_col(aes(fill = type))+
      facet_grid(.~ type, scales = 'free', space = 'free')+
      labs(title = 'Yearly total PM 2.5 Emission in Baltimore City per Source', x ='', y = 'total PM2.5 emissions') +
      theme(legend.position = 'none')

graph3
```

#4
Across united states, how many emissions from coal combustion related sources changed from 1999 to 2008
```{r}
coal_combustion = all_data %>%
      gather(key = 'scc_level', value = 'value', scc_level_one, scc_level_two, scc_level_three, scc_level_four) %>%
      filter(str_detect(value, 'Coal') | str_detect(value, 'Combustion'))

graph4 = coal_combustion %>%
      mutate(year = factor(year)) %>%
      group_by(year) %>%
      summarise(total_emission = sum(emissions, na.rm = TRUE)) %>%
      ggplot(aes(year, total_emission))+
      geom_col(aes(fill = year)) +
      labs(title = 'Total PM2.5 Emissions from Coal or Combustion Sources', x = '', y = 'total PM2.5 emissions')+
      theme(legend.position = 'none')

graph4
```

#5.

Have emissions from motor vehicle sources changed in baltimore?

```{r}
baltimore_vehicles = all_data %>%
      gather(key = 'scc_level', value = 'value', scc_level_one, scc_level_two, scc_level_three, scc_level_four) %>%
      filter(grepl('vehicle',value, ignore.case = TRUE)) %>%
      filter(fips == '24510')

graph5 = baltimore_vehicles %>%
      mutate(year = factor(year)) %>%
      group_by(year) %>%
      summarise(total_emission = sum(emissions, na.rm = TRUE)) %>%
      ggplot(aes(year, total_emission))+
      geom_col(aes(fill = year))+
      labs(title = 'Yearly Total PM2.5 Emission from Vehicles in Baltimore City', 
           x = '',
           y = 'total PM2.5 emissions')+
      theme(legend.position = 'none')

graph5
```


#6
Compare emissions from motor vehicles in baltimore city and los angeles

```{r}
la_baltimore_vehicles = all_data %>%
      gather(key = 'scc_level', value = 'value', scc_level_one, scc_level_two, scc_level_three, scc_level_four) %>%
      filter(grepl('vehicle',value, ignore.case = TRUE)) %>%
      filter(fips == '24510' | fips == '06037')

graph6 = la_baltimore_vehicles %>%
      mutate(year = factor(year),
             fips = factor(fips)) %>%
      group_by(fips, year) %>%
      mutate(fips = ifelse(fips == '24510', 'Baltimore City','LA City')) %>%
      summarise(total_emission = sum(emissions, na.rm = TRUE)) %>%
      ggplot(aes(year, total_emission)) +
      geom_col(aes(fill = fips))+
      facet_grid(.~fips, scales = 'free', space = 'free')+
      labs(title = 'Yearly Total PM2.5 Emission from Vehicles in Baltimore and LA City', x = '', 
           y= 'total PM2.5 emissions') +
      theme(legend.position = 'none')

graph6
```




























